Assessing United States county-level exposure for research on tropical cyclones and human health

### 2020-02-21 writing/anderson_ehp_supp.tex
Make a few edits to main text

\author[1,*]{G. Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1]{Meilin Yan} 
\author[8]{Roger D. Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}

>>>>

\author[1,*]{G. Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1,8]{Meilin Yan} 
\author[9]{Roger D. Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, USA} 
\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, USA} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, USA}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, USA} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, USA}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, USA}
\affil[7]{Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, USA} 
\affil[8]{Beijing Innovation Center for Engineering Science and Advanced Technology 
  and State Key Joint Laboratory of Environmental Simulation and Pollution Control, 
  Peking University, Beijing, China}
\affil[9]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, USA}
\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}

### 2020-02-21 writing/ehp_cover.tex
Make a few edits to main text

\author[1,*]{G.~Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1]{Meilin Yan} 
\author[8]{Roger D.~Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: G.~Brooke Anderson, brooke.anderson@colostate.edu,
  1681 Campus Delivery, Fort Collins, CO, 80523}

>>>>

\author[1,*]{G.~Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1,8]{Meilin Yan} 
\author[9]{Roger D.~Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, USA} 
\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, USA} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, USA}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, USA} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, USA}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, USA}
\affil[7]{Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, USA} 
\affil[8]{Beijing Innovation Center for Engineering Science and Advanced Technology 
  and State Key Joint Laboratory of Environmental Simulation and Pollution Control, 
  Peking University, Beijing, China}
\affil[9]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, USA}
\affil[*]{Corresponding author: G.~Brooke Anderson, brooke.anderson@colostate.edu,
  1681 Campus Delivery, Fort Collins, CO, USA}

### 2020-02-21 writing/ehp_methods.tex
Make a few edits to main text

Two of the exposure metrics (flood- and tornado-based) were inherently binary
in our data, since these metrics were based on whether an event was listed in
the NOAA Storm Events database.  For other exposure metrics, each county
was classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
picked reasonable thresholds, but others could be used with the open-source
data and its associated software.

@@

>>>>

Two of the exposure metrics (flood- and tornado-based) were inherently binary
in our data, since these metrics were based on whether an event was listed in
the NOAA Storm Events database.  For other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
picked reasonable thresholds (e.g., the threshold for gale-force wind for the
wind exposure), but others could be used with the open-source data and its
associated software.

@@

\subsection*{Data and code access}

We have posted study data as an open-source R package
\cite{hurricaneexposuredata}. We have posted additional data and code at
\url{https://github.com/geanders/county_hurricane_exposure}.

### 2020-02-21 writing/ehp_results.tex
Make a few edits to main text

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (when data for all five metrics were available) for
which~100 or more counties were exposed based on at least one metric
(Figure~\ref{fig:jaccard}). In this figure, each row provides results for one
tropical cyclone, and each box in that row shows the Jaccard coefficient for a
pair of metrics.  For almost all tropical cyclones, Jaccard coefficients
were~$\le0.6$ between the distance-based metric and any of the four
hazard-based metrics (i.e., out of the counties assessed as exposed by at least
one of the two metrics in the pair,~60\si{\percent} or fewer were assessed the
same under both metrics). Among the four hazard-based metrics, the
tornado-based metric had universally poor agreement with other metrics in
county-level classification across the tropical cyclones considered.  Agreement
between other pairs of hazard-based metrics were better, but still typically at
best moderate (Jaccard coefficients usually $\le0.5$).

There were, however, exceptions\,---\,tropical cyclones in which exposure
assessment agreed well across several of the metrics considered.  For Floyd
in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
agreed moderately to well for all pairs of exposure metrics except those
involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
in~2003), there was moderate to good agreement for pairwise combinations of
distance, rain, and wind, but poor agreement for other combinations of metrics,
while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
there was moderate to good agreement between distance and rain.

>>>>

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (when data for all five metrics were available) for
which~100 or more counties were exposed based on at least one metric
(Figure~\ref{fig:jaccard}). In this figure, each row provides results for one
tropical cyclone, and each box in that row shows the Jaccard coefficient for a
pair of metrics. For all pairs of metrics, agreement in exposure assessment
was, at best, moderate for all but a few storms. When comparing distance- and
wind-based exposure assessment, only about 10\% of storms had Jaccard indices
higher than 0.6 (i.e., out of the counties assessed as exposed by at least one
of the two metrics in the pair,~60\si{\percent} or more were assessed the same
under both metrics). For comparisons of assessments based on other combinations
of distance-, wind-, rain-, and flood-based metrics, fewer than 5\% of storms
had Jaccard indices above 0.6.  The tornado-based metric had universally poor
agreement with other metrics in county-level classification across the tropical
cyclones considered.

There were a few tropical cyclones in which exposure assessment agreed well
across several of the metrics considered.  For Floyd in~1999 (Figure~S4) and
Irene in~2011, for example, county-level classification agreed moderately to
well for all pairs of exposure metrics except those involving the tornado-based
metric (Figure~\ref{fig:jaccard}). For another set of tropical cyclones (e.g.,
Ernesto in~2006, and Bertha in~1996, Isabel in~2003), there was moderate to
good agreement for pairwise combinations of distance, rain, and wind, but poor
agreement for other combinations of metrics, while for another set of storms
(e.g., Matthew in~2004 and Katrina in~2005), there was moderate to good
agreement between distance and rain.
